Fabric-Elasticity Relationships of Tibial Trabecular Bone are Similar in Osteogenesis Imperfecta and Healthy Individuals

### 2021-10-06 09_Article/01_Revision1_Reviewers/Response to Reviewers.tex

\AR To increase the readability of the table, it was changed in the manuscript, see Table \ref{Table2}.

\begin{sidewaystable*}
	\centering
	\caption{Summary of the tibia ROIs morphological analysis and comparison with literature. Values are presented as mean $\pm$ standard deviation when statistical test is performed on the means or median (inter-quartile range) when test is on medians. The study of Kocijan et al. [16] presents n.s. for non-significant p value test result.}
	\label{Table2}
	\resizebox{\textwidth}{!}{
	\begin{tabular}{cccccccccc}
		\toprule
		\multirow{2}{*}{Variable} & \multirow{2}{*}{Group} & \multicolumn{2}{c}{Present study} & \multicolumn{2}{c}{Folkestad et al. [9]} & \multicolumn{2}{c}{Kocijan et al. [16]} & \multicolumn{2}{c}{Rolvien et al. [26]} \\
		& & Values & p value & Values & p value & Values & p value & Values & p value \\
		\midrule

		\multirow{3}{*}{Age} & Healthy & 41 $\pm$ 14 &  & 54 (21-77) & & 44 (38-52) &  & 49 $\pm$ 16 &  \\
		& \multirow{2}{*}{OI} & \multirow{2}{*}{41 $\pm$ 15} &  & \multirow{2}{*}{53 (21-77)} &  & 42 (35-56)\textsuperscript{1} & & \multirow{2}{*}{46 $\pm$ 16} & \\
		& & &  &  &  & 48 (35-58)\textsuperscript{2} & & & \\[3ex]

		\multirow{3}{*}{BV/TV} & Healthy & 0.222 $\pm$ 0.081 & <0.01 & 0.14 $\pm$ 0.03 & <0.001 & 0.141 (0.130-.0170) & & 0.162 $\pm$ 0.010 & <0.0001 \\
		& \multirow{2}{*}{OI} & \multirow{2}{*}{0.164 $\pm$ 0.079} &  & 0.08 $\pm$ 0.03 &  & 0.098 (0.088-0.114)\textsuperscript{1} & <0.0001\textsuperscript{1} & \multirow{2}{*}{0.095 $\pm$ 0.008} & \\
		& & & & & & 0.081 (0.056-0.092)\textsuperscript{2} & <0.0001\textsuperscript{2} & & \\[3ex]

		\multirow{3}{*}{Tb.N.} & Healthy & 0.842 $\pm$ 0.144 & <0.001 & 1.94 (1.74-2.13) & <0.001 & 1.76 (1.59-2.08) & & 2.143 $\pm$ 0.089 & <0.0001 \\
		& \multirow{2}{*}{OI} & \multirow{2}{*}{0.650 $\pm$ 0.198} &  & 1.30 (0.75-1.53) &  & 1.33 (1.07-1.55)\textsuperscript{1} & <0.0001\textsuperscript{1} & \multirow{2}{*}{1.428 $\pm$ 0.098} & \\
		& & & & & & 0.89 (0.81-1.08)\textsuperscript{2} & <0.001\textsuperscript{2} & & \\[3ex]

		\multirow{3}{*}{Tb.Th.} & Healthy & 0.301 (0.287-0.321) & 0.2 & 0.07 (0.06-0.08) & 0.5 & 0.081 (0.074-0.087) & & 0.075 $\pm$ 0.003 & 0.046 \\
		& \multirow{2}{*}{OI} & \multirow{2}{*}{0.306 (0.292-0.331)} &  & 0.07 (0.06-0.08) &  & 0.074 (0.064-0.090)\textsuperscript{1} & n.s.\textsuperscript{1} & \multirow{2}{*}{0.066 $\pm$ 0.004} & \\
		& & & & & & 0.078 (0.068-0.092)\textsuperscript{2} & n.s.\textsuperscript{2} & & \\[3ex]

		\multirow{3}{*}{Tb.Sp.} & Healthy & 0.924 $\pm$ 0.257 & 0.01 & 0.44 (0.39-0.51) & <0.001 & \multirow{3}{*}{-} & & 0.409 $\pm$ 0.023 & 0.0003 \\
		& \multirow{2}{*}{OI} & \multirow{2}{*}{1.422 $\pm$ 0.694} &  & 0.68 (0.57-1.19) &  & &  & \multirow{2}{*}{0.727 $\pm$ 0.095} & \\
		& & & & & & &  & & \\[3ex]

		\multirow{3}{*}{Tb.Sp.SD} & Healthy & 0.317 $\pm$ 0.136 & 0.02 & 0.20 (0.16-0.25) & <0.001 & 0.221 (0.170-0.242) & & - &  \\
		& \multirow{2}{*}{OI} & \multirow{2}{*}{0.631 $\pm$ 0.383} &  & 0.40 (0.31-1.11) &  & 0.382 (0.311-0.504)\textsuperscript{1} & <0.0001\textsuperscript{1} & - & \\
		& & & & & & 0.698 (0.511-0.890\textsuperscript{2} & <0.001\textsuperscript{2} & & \\[3ex]

		\multirow{3}{*}{SMI} & Healthy & 0.001 (-0.021-0.033) & <0.001 & \multirow{3}{*}{-} &  & \multirow{3}{*}{-} & & \multirow{3}{*}{-} &  \\
		& \multirow{2}{*}{OI} & \multirow{2}{*}{0.056 (0.015-0.079)} & & & & &  & & \\
		& & & & & & & & & \\[3ex]

		\multirow{3}{*}{DA} & Healthy & 1.992 (1.826-2.020) & 0.02 & \multirow{3}{*}{-} & & \multirow{3}{*}{-} & & \multirow{3}{*}{-} &  \\
		& \multirow{2}{*}{OI} & \multirow{2}{*}{2.018 (1.901-2.158)} & & & & & & & \\
		& & & & & & & & & \\[3ex]

		\multirow{3}{*}{ln(CV)} & Healthy & -1.723 $\pm$ 0.344 & <0.0001 & \multirow{3}{*}{-} & & \multirow{3}{*}{-} & & \multirow{3}{*}{-} &  \\
		& \multirow{2}{*}{OI} & \multirow{2}{*}{-1.178 $\pm$ 0.441} & & & & & & & \\
		& & & & & & & & & \\

		\bottomrule
	\end{tabular}}
	OI types included in the study: Present study, types I and IV; Folkestad et al. [9], type I; Kocijan et al. [16], 1) type I 2) type III and IV; Rolvien et al. [26], type I and IV.
\end{sidewaystable*}

### 2021-10-13 09_Article/01_Revision1_Reviewers/Simon_2021_Bone_Manuscript_Revised.tex
Finish Response to Reviewers and manuscript revision

	\begin{abstract}
		Osteogenesis Imperfecta (OI) is an inherited form of bone fragility characterised by impaired synthesis of type I collagen, altered trabecular bone architecture and reduced bone mass. High resolution peripheral computed tomography (HR-pQCT) is a powerful method to investigate bone morphology at peripheral sites including the weight-bearing distal tibia. The resulting 3D reconstructions can be used as a basis of micro-finite element (\si{\micro}FE) or homogenised finite element (hFE) models for bone strength estimation. The hFE scheme uses homogenised local bone volume fraction (BV/TV) and anisotropy information (fabric) to compute healthy bone strength within a reasonable computation time using fabric-elasticity relationships. Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls.

@@

		Compared to healthy controls, we found the OI samples to have significantly lower BV/TV and trabecular number (Tb.N.), significantly higher trabecular separation (Tb.Sp.) and trabecular separation standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.). These results are in agreement with previous studies. The stiffness of ROIs from OI bone reached lower values compared to healthy controls and the multilinear fabric-elasticity fits tended to overestimate the stiffness in the lower range. The filtering of highly heterogeneous ROIs removed these low stiffness ROIs and led to similar correlation coefficients for both OI and healthy groups. Finally, the BV/TV and DA matched data revealed no significant differences in fabric-elasticity parameters between OI and healthy individuals. Compared to previous studies, the stiffness constants from the 61 \si{\micro}m resolution HR-pQCT ROIs were lower than for the 36 \si{\micro}m resolution \si{\micro}CT ROIs.

		In conclusion, despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the mechanical analysis. Since highly heterogenous ROIs coincide with very low BV/TV, we expect them to play a minor role in hFE analysis of distal bone sections.
	\end{abstract}

@@

	High resolution peripheral quantitative computed tomography (HR-pQCT) reconstructions allow \textit{in vivo} assessment of cortical and trabecular architecture and volumetric bone mineral density (BMD) at the distal radius and distal tibia \cite{Boutroy2005}. Moreover, HR-pQCT reconstructions can be used for finite element analysis (FEA) to predict mechanical properties, such as bone stiffness and strength \cite{Boutroy2008}. \Add{HR-pQCT-based FEA becomes more widely used in clinical studies \cite{VanDenBergh2021} and is currently approved by the American Food and Drug Administration (FDA).} Homogenized finite element (hFE) is based on BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort \cite{Pahr2009}. High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,Hosseini2017,AriasMoreno2019}. Thus, it could be legitimate to use hFE on OI affected bone, for estimation of bone strength and potentially associated fracture risk. However, HR-pQCT-based hFE relies on fabric-elasticity relationships. Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.

@@

	\subsection{HR-pQCT}
	HR-pQCT imaging (XtremeCTII, SCANCO Medical, Br\"{u}tisellen,
	Switzerland) was performed at the distal tibia on all patients from both groups. Participants in the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020}. Namely, the reference line was positioned at the proximal margin of the dense structure formed by the tibia plafond and three stacks were scanned proximal to this reference line, see Figure \ref{01_Healthy}. On the other hand, for member of the OI group, the reference line is placed at the subchondral endplate of the ankle joint.  The mid-point of the scanned volume of interest was located at a distance of 7\% of total tibia length from the reference line \cite{GHASEMZADEH2017,Bonaretti2017}, see Figure \ref{01_OI}. Then, the scans were acquired according to the manufacturer's standard in vivo acquisition protocol. The scanned regions according to these two different protocols coincide approximately and are shown side by side in Figure \ref{01_ClinicalSections}. For both the healthy and the OI groups, each stack consisted of 168 slices and an isotropic voxel resolution of 61 \si{\micro}m. This led to a thickness of roughly 10.2 mm for each stack. Standardized scanning settings were used (voltage of 60 kVp, 900 $\mu$A, 100 ms integration time) for the healthy group as well as for the OI group. For the healthy group, motion artefacts of first, middle and last slice (i.e slices number 1, 252, and 504) were graded on a scale proposed by \citeauthor{Pialat2012} \cite{Pialat2012} of 1 (no motion artefacts) to 5 (extreme motion artefacts). The final grade of each scan was defined as the highest slice grade. For the OI group, as the scan consisted of one stack, only one grade was attributed using the same scale as for the control group. Scans were then processed independently of their quality grading. A summary of the scans grading is shown in Figure \ref{01_MotionArtefacts}.

@@

	\subsection{Image analysis}
	The HR-pQCT scans were evaluated using the manufacturer's standard protocol. Briefly, an automatic contouring algorithm was applied to define the periosteal contour of the tibia (masking) and a threshold was applied for segmentation of cortical bone (450 mgHA/cm\textsuperscript{3}) and trabecular bone (320 mgHA/cm\textsuperscript{3}). Then, mask segmented images were used for further analysis.\\

@@

	A \si{\micro}FE analysis was performed using \textsc{ABAQUS 6.14}. In brief, each voxel of the cleaned ROI was converted to a fully integrated linear brick elements (C3D8) using a direct voxel conversion approach. Then, a Young's modulus $E$ of 10,000 MPa and a Poisson's ratio $\nu$ of 0.3 were assigned to each element. The homogenisation process consisted of 6 independent simulations of different load cases, 3 uni-axial and 3 simple shear cases, using kinematic uniform boundary conditions (KUBCs) \cite{Panyasantisuk2015}. Unlike periodicity-compatible mixed uniform boundary conditions (PMUBCs), KUBCs do not require to rotate the ROI into the respective fabric coordinate system. Such rotation would potentially decrease image quality. The homogenization process allowed to calculate the components of the stiffness tensor and to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}. This model computes the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV or $\rho$, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$, as shown in Equation \ref{Eq202}.\\

@@

	\subsection{Linear Regression}
	The orthotropic stiffness tensors obtained after transformation onto fabric coordinate system were then used to perform a multiple linear regression with the Zysset-Curnier model. Standard linear models assume independent and identically distributed (iid) variables. As this assumption was violated by the fact that six ROIs were analysed per individual, a linear mixed-effect model was preferred. This last model, shown in Equation \ref{Eq204} in Laird-Ware form \cite{Laird1982}, considered the non-independence of ROIs from the same individual. A more detailed form of this model is presented in Appendix \ref{A1}.

@@

	\section{Results}
	%
	\subsection{Morphological Analysis}
	The results of the morphological analysis are summarised in Table \ref{Table2}. The individual matching for age and sex allowed to have similar group distributions with 17 females and 11 males in each group. The mean age was 41y $\pm$ 14y and 41y $\pm$ 15y for the mat\-ched healthy and OI individuals, respectively. BV/TV of healthy individuals was higher than BV/TV of OI group with a difference 95\% CI of [0.016, 0.101] and p-value <0.01. Similarly, trabecular number was higher in the matched healthy group compared to the matched OI group with a difference 95\% CI of [0.099, 0.285] and a corresponding p-value <0.001. The trabecular thickness does not show significant differences between groups. On the other hand, permutation test performed for trabecular separation showed that it is higher in matched OI group compared to healthy individuals with a p-value of 0.01 and an exclusion range of ($-\infty$, -0.384] $\cup$ [0.421, $\infty$). Trabecular separation SD is higher in OI individuals compared to matched healthy individuals with a p-value of 0.02 and an exclusion range of ($-\infty$, -0.232] $\cup$ [0.251, $\infty$). SMI as well as the degree of anisotropy are higher for matched OI than for healthy people with p-values <0.001 and of 0.02, respectively. Finally, the log transformation of the coefficient of variation gives the stronger difference in means with a p value <0.0001 and a 95\% CI of [-0.757, -0.333] where CV is higher in matched OI individuals compared to matched healthy individuals.\\

@@

	\begin{sidewaystable*}
		\centering
		\caption{Summary of the tibia ROIs morphological analysis and comparison with literature. Values are presented as mean $\pm$ standard deviation when statistical test is performed on the means or median (inter-quartile range) when test is on medians. The study of \cite{Kocijan2015} presents n.s. for non-significant p value test result.}
		\label{Table2}
		\begin{tabular}{cccccccccc}
			\toprule
			\multirow{2}{*}{Variable} & \multirow{2}{*}{Group} & \multicolumn{2}{c}{Present study} & \multicolumn{2}{c}{\citeauthor{Folkestad2012}\cite{Folkestad2012}} & \multicolumn{2}{c}{\citeauthor{Kocijan2015}\cite{Kocijan2015}} & \multicolumn{2}{c}{\citeauthor{Rolvien2018}\cite{Rolvien2018}} \\
			& & Values & p value & Values & p value & Values & p value & Values & p value \\
			\midrule

			\multirow{3}{*}{Age} & Healthy & 41 $\pm$ 14 &  & 54 (21-77) & & 44 (38-52) &  & 49 $\pm$ 16 &  \\
			& OI Type I & \multirow{2}{*}{41 $\pm$ 15} &  & 53 (21-77) &  & 42 (35-56) & & \multirow{2}{*}{46 $\pm$ 16} & \\
			&  OI Type III \& IV & &  &  &  & 48 (35-58) & & & \\[3ex]

			\multirow{3}{*}{BV/TV} & Healthy & 0.222 $\pm$ 0.081 & <0.01 & 0.14 $\pm$ 0.03 & <0.001 & 0.141 (0.130-.0170) & & 0.162 $\pm$ 0.010 & <0.0001 \\
			& OI Type I & \multirow{2}{*}{0.164 $\pm$ 0.079} &  & 0.08 $\pm$ 0.03 &  & 0.098 (0.088-0.114) & <0.0001 & \multirow{2}{*}{0.095 $\pm$ 0.008} & \\
			&  OI Type III \& IV & & & & & 0.081 (0.056-0.092) & <0.0001 & & \\[3ex]

			\multirow{3}{*}{Tb.N.} & Healthy & 0.842 $\pm$ 0.144 & <0.001 & 1.94 (1.74-2.13) & <0.001 & 1.76 (1.59-2.08) & & 2.143 $\pm$ 0.089 & <0.0001 \\
			& OI Type I & \multirow{2}{*}{0.650 $\pm$ 0.198} &  & 1.30 (0.75-1.53) &  & 1.33 (1.07-1.55) & <0.0001 & \multirow{2}{*}{1.428 $\pm$ 0.098} & \\
			&  OI Type III \& IV & & & & & 0.89 (0.81-1.08) & <0.001 & & \\[3ex]

			\multirow{3}{*}{Tb.Th.} & Healthy & 0.301 (0.287-0.321) & 0.2 & 0.07 (0.06-0.08) & 0.5 & 0.081 (0.074-0.087) & & 0.075 $\pm$ 0.003 & 0.046 \\
			& OI Type I & \multirow{2}{*}{0.306 (0.292-0.331)} &  & 0.07 (0.06-0.08) &  & 0.074 (0.064-0.090) & n.s. & \multirow{2}{*}{0.066 $\pm$ 0.004} & \\
			&  OI Type III \& IV & & & & & 0.078 (0.068-0.092) & n.s. & & \\[3ex]

			\multirow{3}{*}{Tb.Sp.} & Healthy & 0.924 $\pm$ 0.257 & 0.01 & 0.44 (0.39-0.51) & <0.001 & - & & 0.409 $\pm$ 0.023 & 0.0003 \\
			& OI Type I & \multirow{2}{*}{1.422 $\pm$ 0.694} &  & 0.68 (0.57-1.19) &  & - &  & \multirow{2}{*}{0.727 $\pm$ 0.095} & \\
			&  OI Type III \& IV & & & & & - &  & & \\[3ex]

			\multirow{3}{*}{Tb.Sp.SD} & Healthy & 0.317 $\pm$ 0.136 & 0.02 & 0.20 (0.16-0.25) & <0.001 & 0.221 (0.170-0.242) & & - &  \\
			& OI Type I & \multirow{2}{*}{0.631 $\pm$ 0.383} &  & 0.40 (0.31-1.11) &  & 0.382 (0.311-0.504) & <0.0001 & - & \\
			&  OI Type III \& IV & & & & & 0.698 (0.511-0.890) & <0.001 & & \\[3ex]

			\multirow{3}{*}{SMI} & Healthy & 0.001 (-0.021-0.033) & <0.001 & - &  & - & & - &  \\
			& OI Type I & \multirow{2}{*}{0.056 (0.015-0.079)} &  & - &  & - &  & - & \\
			&  OI Type III \& IV & & & & & - &  & & \\[3ex]

			\multirow{3}{*}{DA} & Healthy & 1.992 (1.826-2.020) & 0.02 & - &  & - & & - &  \\
			& OI Type I & \multirow{2}{*}{2.018 (1.901-2.158)} &  & - &  & - &  & - & \\
			&  OI Type III \& IV & & & & & - &  & & \\[3ex]

			\multirow{3}{*}{ln(CV)} & Healthy & -1.723 $\pm$ 0.344 & <0.0001 & - &  & - & & - &  \\
			& OI Type I & \multirow{2}{*}{-1.178 $\pm$ 0.441} &  & - &  & - &  & - & \\
			&  OI Type III \& IV & & & & & - &  & & \\

			\bottomrule
		\end{tabular}
	\end{sidewaystable*}

>>>>

	\begin{abstract}
		Osteogenesis Imperfecta (OI) is an inherited form of bone fragility characterised by impaired synthesis of type I collagen, altered trabecular bone architecture and reduced bone mass. High resolution peripheral computed tomography (HR-pQCT) is a powerful method to investigate bone morphology at peripheral sites including the weight-bearing distal tibia. The resulting 3D reconstructions can be used as a basis of micro-finite element (\si{\micro}FE) or homogenised finite element (hFE) models for bone strength estimation. The hFE scheme uses homogenised local bone volume fraction (BV/TV) and anisotropy information (fabric) to compute healthy bone strength within a reasonable computation time using fabric-elasticity relationships. \Add{However, it is unclear if these relationships quantified previously for healthy controls are valid for trabecular bone from OI patients.} Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls.

@@

		Compared to healthy controls, we found the OI samples to have significantly lower \Add{CV,} BV/TV and trabecular number (Tb.N.), significantly higher trabecular separation (Tb.Sp.) and trabecular separation standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.). These results are in agreement with previous studies. \Del{The stiffness of ROIs from OI bone reached lower values compared to healthy controls and the multilinear fabric-elasticity fits tended to overestimate the stiffness in the lower range}. \Add{The stiffnesses of highly heterogeneous ROIs are randomly lower with respect to the fabric-elasticity relationships and reflects the limit of validity of the computational homogenisation methodology.}

		The filtering of highly heterogeneous ROIs removed these low stiffness ROIs and led to similar correlation coefficients for both OI and healthy groups. Finally, the BV/TV and DA matched data revealed no significant differences in fabric-elasticity parameters between OI and healthy individuals. Compared to previous studies, the stiffness constants from the 61 \si{\micro}m resolution HR-pQCT ROIs were lower than for the 36 \si{\micro}m resolution \si{\micro}CT ROIs.

		In conclusion, \Add{OI trabecular bone of the distal tibia was shown to be significantly more heterogeneous and less dense than healthy controls.} Despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the computational stiffness analysis. \Del{Since highly heterogenous ROIs coincide with low BV/TV and can be handled smoothly by the fabric-elasticity relationships we expect them to play a minor role in hFE analysis of distal bone sections.} \Add{Accordingly, the elastic properties used for FEA of healthy bones are also valid for OI bones.} 
	\end{abstract}

@@

	High resolution peripheral quantitative computed tomography (HR-pQCT) reconstructions allow \textit{in vivo} assessment of cortical and trabecular architecture and volumetric bone mineral density (BMD) at the distal radius and distal tibia \cite{Boutroy2005}. Moreover, HR-pQCT reconstructions can be used for finite element analysis (FEA) to predict mechanical properties, such as bone stiffness and strength \cite{Boutroy2008} \Add{used for fracture risk estimation}. \Add{Micro-finite element (\si{\micro}FE) analysis converts each voxel into an element and thus requires high performance computer for a bone section analysis}. Homogenized finite element (hFE) is based on \Add{local} BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort \cite{Pahr2009}. High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,Hosseini2017,AriasMoreno2019}. \Add{Homogenised QCT-based FEA has become more widely used in clinical studies \cite{VanDenBergh2021} and is currently approved by the American Food and Drug Administration (FDA).} Thus, it could be legitimate to use hFE on OI affected bone, for estimation of bone strength and potentially associated fracture risk. However, HR-pQCT-based hFE relies on fabric-elasticity relationships. Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.\Add{ Similar fabric-elasticity relationships will allow to extend the application of HR-pQCT-based hFE of healthy bone to OI bone at least in the linear elastic regime.}

@@

	\subsection{HR-pQCT}
	HR-pQCT imaging (XtremeCTII, SCANCO Medical, Br\"{u}tisellen,
	Switzerland) was performed at the distal tibia on all patients from both groups. \Add{The distal tibia was chosen because numerous studies were performed on the distal radius \cite{AriasMoreno2019,Boutroy2008,Gross2013,Hosseini2017,Kocijan2015,Varga2011}, some study performed analysis on both distal radius and tibia with XtremeCTI (XtremeCT) \cite{Boutroy2005,Folkestad2012} but only one study was performed on the distal tibia with XtremeCTII \cite{Schenk2020}.} Participants in the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020}. Namely, the reference line was positioned at the proximal margin of the dense structure formed by the tibia plafond and three stacks were scanned proximal to this reference line, see Figure \ref{01_Healthy}. On the other hand, for member of the OI group, the reference line is placed at the subchondral endplate of the ankle joint.  The mid-point of the scanned volume of interest was located at a distance of 7\% of total tibia length from the reference line \cite{GHASEMZADEH2017,Bonaretti2017}, see Figure \ref{01_OI}. Then, the scans were acquired according to the manufacturer's standard in vivo acquisition protocol. The scanned regions according to these two different protocols coincide approximately and are shown side by side in Figure \ref{01_ClinicalSections}. For both the healthy and the OI groups, each stack consisted of 168 slices and an isotropic voxel resolution of 61 \si{\micro}m. This led to a thickness of roughly 10.2 mm for each stack. Standardized scanning settings were used (voltage of 60 kVp, 900 $\mu$A, 100 ms integration time) for the healthy group as well as for the OI group. For the healthy group, motion artefacts of first, middle and last slice (i.e slices number 1, 252, and 504) were graded on a scale proposed by \citeauthor{Pialat2012} \cite{Pialat2012} of 1 (no motion artefacts) to 5 (extreme motion artefacts). The final grade of each scan was defined as the highest slice grade. For the OI group, as the scan consisted of one stack, only one grade was attributed using the same scale as for the control group. Scans were then processed independently of their quality grading. A summary of the scans grading is shown in Figure \ref{01_MotionArtefacts}.

@@

	\subsection{Image analysis}
	The HR-pQCT scans were evaluated using the manufacturer's standard protocol. Briefly, an automatic contouring algorithm was applied to define the periosteal contour of the tibia (masking)\Add{. Then, a Gauss filtering (sigma = 0.8, support = 1 voxel) was performed} and a threshold was applied for segmentation of cortical bone (450 mgHA/cm\textsuperscript{3}) and trabecular bone (320 mgHA/cm\textsuperscript{3}). Then, mask segmented images were used for further analysis.\\

@@

	A \si{\micro}FE analysis was performed using \textsc{ABAQUS 6.14}. In brief, each voxel of the cleaned ROI was converted to a fully integrated linear brick elements (C3D8) using a direct voxel conversion approach. Then, a Young's modulus $E$ of 10,000 MPa and a Poisson's ratio $\nu$ of 0.3 were assigned to each element. The homogenisation process consisted of 6 independent simulations of different load cases, 3 uni-axial and 3 simple shear cases, using kinematic uniform boundary conditions (KUBCs) \cite{Panyasantisuk2015}. Unlike periodicity-compatible mixed uniform boundary conditions (PMUBCs), KUBCs do not require to rotate the ROI into the respective fabric coordinate system. Such rotation would potentially decrease image quality \Add{due to interpolation}. The homogenization process allowed to calculate the components of the stiffness tensor and to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}. This model computes the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV or $\rho$, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$, as shown in Equation \ref{Eq202}.\\

@@

	\subsection{Linear Regression}
	The orthotropic stiffness tensors obtained after transformation onto fabric coordinate system were then used to perform a multiple linear regression with the Zysset-Curnier model \Add{using the entire sample of both groups}. Standard linear models assume independent and identically distributed (iid) variables. As this assumption was violated by the fact that six ROIs were analysed per individual, a linear mixed-effect model was preferred. This last model, shown in Equation \ref{Eq204} in Laird-Ware form \cite{Laird1982}, considered the non-independence of ROIs from the same individual. A more detailed form of this model is presented in Appendix \ref{A1}.

@@

	\section{Results}
	%
	\subsection{Morphological Analysis}
	The results of the morphological analysis are summarised in Table \ref{Table2}. The individual matching for age and sex allowed to have similar group distributions with 17 females and 11 males in each group. The mean age was 41y $\pm$ 14y and 41y $\pm$ 15y for the mat\-ched healthy and OI individuals, respectively. \Add{The OI group matched with control was composed of nineteen type I and nine type IV.} BV/TV of healthy individuals was higher than BV/TV of OI group with a difference 95\% CI of [0.016, 0.101] and p-value <0.01. Similarly, trabecular number was higher in the matched healthy group compared to the matched OI group with a difference 95\% CI of [0.099, 0.285] and a corresponding p-value <0.001. The trabecular thickness does not show significant differences between groups. On the other hand, permutation test performed for trabecular separation showed that it is higher in matched OI group compared to healthy individuals with a p-value of 0.01 and an exclusion range of ($-\infty$, -0.384] $\cup$ [0.421, $\infty$). Trabecular separation SD is higher in OI individuals compared to matched healthy individuals with a p-value of 0.02 and an exclusion range of ($-\infty$, -0.232] $\cup$ [0.251, $\infty$). SMI as well as the degree of anisotropy are higher for matched OI than for healthy people with p-values <0.001 and of 0.02, respectively. Finally, the log transformation of the coefficient of variation gives the stronger difference in means with a p value <0.0001 and a 95\% CI of [-0.757, -0.333] where CV is higher in matched OI individuals compared to matched healthy individuals.\\

@@

	\begin{sidewaystable*}
		\centering
		\caption{Summary of the tibia ROIs morphological analysis and comparison with literature. Values are presented as mean $\pm$ standard deviation when statistical test is performed on the means or median (inter-quartile range) when test is on medians. The study of Kocijan et al. [16] presents n.s. for non-significant p value test result.}
		\label{Table2}
		\resizebox{\textwidth}{!}{
			\begin{tabular}{cccccccccc}
				\toprule
				\multirow{2}{*}{Variable} & \multirow{2}{*}{Group} & \multicolumn{2}{c}{Present study} & \multicolumn{2}{c}{\citeauthor{Folkestad2012}\cite{Folkestad2012}} & \multicolumn{2}{c}{\citeauthor{Kocijan2015}\cite{Kocijan2015}} & \multicolumn{2}{c}{\citeauthor{Rolvien2018}\cite{Rolvien2018}} \\
				& & Values & p value & Values & p value & Values & p value & Values & p value \\
				\midrule

				\multirow{3}{*}{Age} & Healthy & 41 $\pm$ 14 &  & 54 (21-77) & & 44 (38-52) &  & 49 $\pm$ 16 &  \\
				& \multirow{2}{*}{OI} & \multirow{2}{*}{41 $\pm$ 15} &  & \multirow{2}{*}{53 (21-77)} &  & 42 (35-56)\textsuperscript{1} & & \multirow{2}{*}{46 $\pm$ 16} & \\
				& & &  &  &  & 48 (35-58)\textsuperscript{2} & & & \\[3ex]

				\multirow{3}{*}{BV/TV} & Healthy & 0.222 $\pm$ 0.081 & & 0.14 $\pm$ 0.03 & & 0.141 (0.130-.0170) & & 0.162 $\pm$ 0.010 & \\
				& \multirow{2}{*}{OI} & \multirow{2}{*}{0.164 $\pm$ 0.079} & \multirow{2}{*}{<0.01} & \multirow{2}{*}{0.08 $\pm$ 0.03} & \multirow{2}{*}{<0.001} & 0.098 (0.088-0.114)\textsuperscript{1} & <0.0001\textsuperscript{1} & \multirow{2}{*}{0.095 $\pm$ 0.008} & \multirow{2}{*}{<0.0001} \\
				& & & & & & 0.081 (0.056-0.092)\textsuperscript{2} & <0.0001\textsuperscript{2} & & \\[3ex]

				\multirow{3}{*}{Tb.N.} & Healthy & 0.842 $\pm$ 0.144 & & 1.94 (1.74-2.13) & & 1.76 (1.59-2.08) & & 2.143 $\pm$ 0.089 & \\
				& \multirow{2}{*}{OI} & \multirow{2}{*}{0.650 $\pm$ 0.198} & \multirow{2}{*}{<0.001} & \multirow{2}{*}{1.30 (0.75-1.53)} & \multirow{2}{*}{<0.001} & 1.33 (1.07-1.55)\textsuperscript{1} & <0.0001\textsuperscript{1} & \multirow{2}{*}{1.428 $\pm$ 0.098} & \multirow{2}{*}{<0.0001} \\
				& & & & & & 0.89 (0.81-1.08)\textsuperscript{2} & <0.001\textsuperscript{2} & & \\[3ex]

				\multirow{3}{*}{Tb.Th.} & Healthy & 0.301 (0.287-0.321) & & 0.07 (0.06-0.08) & & 0.081 (0.074-0.087) & & 0.075 $\pm$ 0.003 & \\
				& \multirow{2}{*}{OI} & \multirow{2}{*}{0.306 (0.292-0.331)} & \multirow{2}{*}{0.2} & \multirow{2}{*}{0.07 (0.06-0.08)} & \multirow{2}{*}{0.5} & 0.074 (0.064-0.090)\textsuperscript{1} & n.s.\textsuperscript{1} & \multirow{2}{*}{0.066 $\pm$ 0.004} & \multirow{2}{*}{0.046}\\
				& & & & & & 0.078 (0.068-0.092)\textsuperscript{2} & n.s.\textsuperscript{2} & & \\[3ex]

				\multirow{3}{*}{Tb.Sp.} & Healthy & 0.924 $\pm$ 0.257 & & 0.44 (0.39-0.51) & & \multirow{3}{*}{-} & & 0.409 $\pm$ 0.023 & \\
				& \multirow{2}{*}{OI} & \multirow{2}{*}{1.422 $\pm$ 0.694} & \multirow{2}{*}{0.01} & \multirow{2}{*}{0.68 (0.57-1.19)} & \multirow{2}{*}{<0.001} & &  & \multirow{2}{*}{0.727 $\pm$ 0.095} & \multirow{2}{*}{0.0003}\\
				& & & & & & & & & \\[3ex]

				\multirow{3}{*}{Tb.Sp.SD} & Healthy & 0.317 $\pm$ 0.136 & & 0.20 (0.16-0.25) & & 0.221 (0.170-0.242) & & - &  \\
				& \multirow{2}{*}{OI} & \multirow{2}{*}{0.631 $\pm$ 0.383} & \multirow{2}{*}{0.02} & \multirow{2}{*}{0.40 (0.31-1.11)} & \multirow{2}{*}{<0.001} & 0.382 (0.311-0.504)\textsuperscript{1} & <0.0001\textsuperscript{1} & - & \\
				& & & & & & 0.698 (0.511-0.890\textsuperscript{2} & <0.001\textsuperscript{2} & & \\[3ex]

				\multirow{3}{*}{SMI} & Healthy & 0.001 (-0.021-0.033) & & \multirow{3}{*}{-} &  & \multirow{3}{*}{-} & & \multirow{3}{*}{-} &  \\
				& \multirow{2}{*}{OI} & \multirow{2}{*}{0.056 (0.015-0.079)} & \multirow{2}{*}{<0.001} & & & &  & & \\
				& & & & & & & & & \\[3ex]
